---
title: "Exploratory Analysis"
author: "Zaynub Ibrahim (zi2125)"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(readr)
library(plotly)

knitr::opts_chunk$set(
	fig.width = 9, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Clean data:

  * Clean names
  * Select relevant predictors & outcomes
  * Drop neighborhoods that have no population
  
```{r}
plots_df = read_excel("./data/nta-metadata.xlsx", sheet = "NTA Data") %>%
  janitor::clean_names() %>%
  select(nta_name, nta_code, total_pop, hispanic:other_race, poverty, smm, gonorrhea, health_ins, edu_less_than_hs, preterm_births, late_no_prenatal_care, medicaid_enroll) %>%
  drop_na(total_pop) %>% 
    mutate(poverty_level = cut(poverty, breaks = c(-Inf, 10, 20, 30, 40, Inf), labels = c("poverty_10","poverty_20", "poverty_30", "poverty_40", "poverty_40+"))) %>% 
  pivot_longer(
   cols = hispanic:other_race,
   names_to = "race",
   values_to = "percent_pop",
   values_drop_na = TRUE
 ) 
```


### Outcome: SMM

poverty Vs SMM 
```{r}
poverty_smm_ggplot = 
  plots_df %>% 
  ggplot(aes(x = poverty, y = smm), group = nta_name) +
  geom_point(color = "red")

ggplotly(poverty_smm_ggplot)



poverty2_smm_ggplot = 
  plots_df %>% 
  ggplot(aes(x = poverty_level, y = smm)) + 
  geom_boxplot()

ggplotly(poverty2_smm_ggplot)


```

late or no prenatal care vs SMM
```{r}
prenatal_care_ggplot = 
  plots_df %>% 
  ggplot(aes(x = late_no_prenatal_care, y = smm, group = nta_name)) + 
  geom_point(color = "red")

ggplotly(prenatal_care_ggplot)
```


### Outcome: Gonorrhea

gonorrhea vs health insurance 
```{r}
gonorrhea1_ggplot = 
  plots_df %>% 
  ggplot(aes(x = health_ins, y = gonorrhea), group = nta_name) + 
  geom_point(color = "green")

ggplotly(gonorrhea1_ggplot)

```

gonorrhea vs medicaid 
```{r}
gonorrhea2_ggplot = 
  plots_df %>% 
  ggplot(aes(x = medicaid_enroll, y = gonorrhea), group = nta_name) +
  geom_point(color = "green")

ggplotly(gonorrhea2_ggplot)

```

gonorrhea vs education level 
```{r}
gonorrhea3_ggplot = 
  plots_df %>% 
  ggplot(aes(x = edu_less_than_hs, y = gonorrhea), group = nta_name) +
  geom_point(color = "green") 


ggplotly(gonorrhea3_ggplot)

```


### Outcome: Preterm Births

health insurance vs preterm birth
```{r}
preterm_ggplot = 
  plots_df %>% 
  ggplot(aes(x = health_ins, y = preterm_births), group = nta_name) +
  geom_point(color = "blue")

ggplotly(preterm_ggplot)
```

medicaid enrollment vs preterm birth
```{r}
preterm2_ggplot = 
  plots_df %>% 
  ggplot(aes(x = medicaid_enroll, y = preterm_births, group = nta_name)) + 
  geom_point(color = "blue")

ggplotly(preterm2_ggplot)
```

late or no prenatal care vs preterm births 
```{r}
prenatal_care_ggplot = 
  plots_df %>% 
  ggplot(aes(x = late_no_prenatal_care, y = preterm_births, group = nta_name)) + 
  geom_point(color = "blue")

ggplotly(prenatal_care_ggplot)
```





